Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-05-15 AppendixI.Rmd
Results & Appendix

---
title: "Appendix I"
author:
  - name: Ariane MIRABEL
    authsuperscript: 1*
  - name: Eric MARCON
    authsuperscript: 1
  - name: Bruno HERAULT
    authsuperscript: 2
affiliation:
  - affsuperscript: 1
    dptuniv: UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles, Université de Guyane.
    address: >
      Campus Agronomique,
      97310 Kourou,
      France.
  - affsuperscript: 2
    dptuniv: INPHB (Institut National Ploytechnique Félix Houphoüet Boigny) 
    address: Yamoussoukro, Ivory Coast
corrauthor:
    email: ariane.mirabel@ecofog.gf
    url: http://www.ecofog.gf/spip.php?article47
journalinfo: Hal 00679993
date: "`r format(Sys.time(), '%d %B %Y')`"
preamble: >
  \usepackage{textcomp}
  \DeclareUnicodeCharacter{B0}{\textdegree}
  \usepackage{tabu}
  \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
  \hyphenation{bio-di-ver-si-ty sap-lings}
lang: english
output: 
  bookdown::pdf_book:
    base_format: EcoFoG::article
    keep_tex: TRUE
  bookdown::gitbook: default
  bookdown::word_document2: default
---

@@

**Figure A3**: Trajectories of the Rao functional diversity over 30 years after disturbance. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes. \@ref(fig:DivFunAppendix).

**Figure A4**: Trajectories of the Rao functional diversity over 30 years after disturbance. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes. \@ref(fig:PCA_graphs).

**Figure A5**: Trajectories of the Rao functional diversity over 30 years after disturbance. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes. \@ref(fig:RedFun).

@@

# **Figure A3**: Rao diversity trajectories with all plots display

```{r DivFunAppendix, out.width = '60%', echo=FALSE, fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.", fig.env="figure*"}
load("DB/FunctionalTraj_ForGraphs")

@@

# **Figure A4**: Functional PCA

@@

# **Figure A5**: Overall functional redundancy

>>>>

---
title: "Appendix I"
author:
  - name: Ariane MIRABEL
    authsuperscript: 1*
  - name: Eric MARCON
    authsuperscript: 1
  - name: Bruno HERAULT
    authsuperscript: 2
affiliation:
  - affsuperscript: 1
    dptuniv: UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles, Université de Guyane.
    address: >
      Campus Agronomique,
      97310 Kourou,
      France.
  - affsuperscript: 2
    dptuniv: INPHB (Institut National Ploytechnique Félix Houphoüet Boigny) 
    address: Yamoussoukro, Ivory Coast
corrauthor:
    email: ariane.mirabel@ecofog.gf
    url: http://www.ecofog.gf/spip.php?article47
journalinfo: Hal 00679993
date: "`r format(Sys.time(), '%d %B %Y')`"
preamble: >
  \usepackage{textcomp}
  \DeclareUnicodeCharacter{B0}{\textdegree}
  \usepackage{tabu}
  \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
  \hyphenation{bio-di-ver-si-ty sap-lings}
lang: english
output: 
  bookdown::gitbook: default
  bookdown::pdf_book:
    base_format: EcoFoG::article
    keep_tex: TRUE
  bookdown::word_document2: default
---

@@

**Figure A3**: Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversities. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3 \@ref(fig:DivTaxo).

**Figure A4**: Trajectories of the Rao functional diversity over 30 years after disturbance. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes. \@ref(fig:DivFun).

**Figure A5**: Summarizing plots of the functional PCA analysis defining communities functional space. **(a)** Plot of the samples from the BRIDGE functional database in the two dimensional main plan. **(b)** bow plot of the PCA eignevalues in percentage. **(c)** Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem traits, and maximum species height). \@ref(fig:PCA_graphs).

**Figure A6**: Trajectories of the functional redundancy over 30 years after disturbance. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3. \@ref(fig:RedFun).

@@

# **Figure A3**: Richness, Shannon and Simpson taxonomic diversity trajectories

```{r DivTaxo, out.width = '100%', fig.asp = 1.5/3, echo=FALSE, fig.cap="Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities (ref:Bolda) Richness, (ref:Boldb) Shannon and (ref:Boldc) Simpson diversities. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
load("DB/ReplacementTraj_ForGraphs")

TaxoTraj(CompleteTaxo)

# **Figure A4**: Rao diversity trajectories with all plots display

```{r DivFun, out.width = '60%', echo=FALSE, fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.", fig.env="figure*"}
load("DB/FunctionalTraj_ForGraphs")

@@

# **Figure A5**: Functional PCA

@@

# **Figure A6**: Overall functional redundancy

### 2018-05-15 AppendixI.tex
Results & Appendix

\textbf{Figure A3}: Trajectories of the Rao functional diversity over 30
years after disturbance. Trajectories correspond to the median (solid
line) and 0.025 and 0.975 percentile (gray envelope) observed after 50
iteration of the taxonomic uncertainty propagation. Initial treatments
are represented by solid lines colorswith green for control, blue for
T1,orange for T2 and red for T3.and the missing trait value filling
processes. \ref{fig:DivFunAppendix}.

\textbf{Figure A4}: Trajectories of the Rao functional diversity over 30
years after disturbance. Trajectories correspond to the median (solid
line) and 0.025 and 0.975 percentile (gray envelope) observed after 50
iteration of the taxonomic uncertainty propagation. Initial treatments
are represented by solid lines colorswith green for control, blue for
T1,orange for T2 and red for T3.and the missing trait value filling
processes. @ref(fig:PCA\_graphs).

\textbf{Figure A5}: Trajectories of the Rao functional diversity over 30
years after disturbance. Trajectories correspond to the median (solid
line) and 0.025 and 0.975 percentile (gray envelope) observed after 50
iteration of the taxonomic uncertainty propagation. Initial treatments
are represented by solid lines colorswith green for control, blue for
T1,orange for T2 and red for T3.and the missing trait value filling
processes. \ref{fig:RedFun}.

@@

\section{\texorpdfstring{\textbf{Figure A3}: Rao diversity trajectories
with all plots
display}{Figure A3: Rao diversity trajectories with all plots display}}\label{figure-a3-rao-diversity-trajectories-with-all-plots-display}

@@

{\centering \includegraphics[width=0.6\linewidth]{AppendixI_files/figure-latex/DivFunAppendix-1}

@@

\caption{Trajectories of the Rao functional diversity over 30 years after disturbance. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.}\label{fig:DivFunAppendix}
\end{figure*}

\section{\texorpdfstring{\textbf{Figure A4}: Functional
PCA}{Figure A4: Functional PCA}}\label{figure-a4-functional-pca}

@@

\section{\texorpdfstring{\textbf{Figure A5}: Overall functional
redundancy}{Figure A5: Overall functional redundancy}}\label{figure-a5-overall-functional-redundancy}

>>>>

\textbf{Figure A3}: Trajectories of the difference to the 1989
inventories (5 years after disturbance) over 30 years after disturbance
of plots communities \textbf{(a)} Richness, \textbf{(b)} Shannon and
\textbf{(c)} Simpson diversities. Trajectories correspond to the median
(solid line) and 0.025 and 0.975 percentile (gray envelope) observed
after 50 iteration of the taxonomic uncertainty propagation. Initial
treatments are represented by solid lines colors with green for control,
blue for T1,orange for T2 and red for T3 \ref{fig:DivTaxo}.

\textbf{Figure A4}: Trajectories of the Rao functional diversity over 30
years after disturbance. Trajectories correspond to the median (solid
line) and 0.025 and 0.975 percentile (gray envelope) observed after 50
iteration of the taxonomic uncertainty propagation. Initial treatments
are represented by solid lines colorswith green for control, blue for
T1,orange for T2 and red for T3.and the missing trait value filling
processes. \ref{fig:DivFun}.

\textbf{Figure A5}: Summarizing plots of the functional PCA analysis
defining communities functional space. \textbf{(a)} Plot of the samples
from the BRIDGE functional database in the two dimensional main plan.
\textbf{(b)} bow plot of the PCA eignevalues in percentage. \textbf{(c)}
Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem
traits, and maximum species height). @ref(fig:PCA\_graphs).

\textbf{Figure A6}: Trajectories of the functional redundancy over 30
years after disturbance. Trajectories correspond to the median (solid
line) and 0.025 and 0.975 percentile (gray envelope) observed after 50
iteration of the taxonomic uncertainty propagation and the missing trait
value filling processes. Initial treatments are represented by solid
lines colors with green for control, blue for T1,orange for T2 and red
for T3. \ref{fig:RedFun}.

@@

\section{\texorpdfstring{\textbf{Figure A3}: Richness, Shannon and
Simpson taxonomic diversity
trajectories}{Figure A3: Richness, Shannon and Simpson taxonomic diversity trajectories}}\label{figure-a3-richness-shannon-and-simpson-taxonomic-diversity-trajectories}

@@

{\centering \includegraphics[width=1\linewidth]{AppendixI_files/figure-latex/DivTaxo-1}

@@

\caption{Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities \textbf{(a)} Richness, \textbf{(b)} Shannon and \textbf{(c)} Simpson diversities. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:DivTaxo}
\end{figure*}

\section{\texorpdfstring{\textbf{Figure A4}: Rao diversity trajectories
with all plots
display}{Figure A4: Rao diversity trajectories with all plots display}}\label{figure-a4-rao-diversity-trajectories-with-all-plots-display}

{\centering \includegraphics[width=0.6\linewidth]{AppendixI_files/figure-latex/DivFun-1}

\caption{Trajectories of the Rao functional diversity over 30 years after disturbance. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.}\label{fig:DivFun}
\end{figure*}

\section{\texorpdfstring{\textbf{Figure A5}: Functional
PCA}{Figure A5: Functional PCA}}\label{figure-a5-functional-pca}

@@

\section{\texorpdfstring{\textbf{Figure A6}: Overall functional
redundancy}{Figure A6: Overall functional redundancy}}\label{figure-a6-overall-functional-redundancy}

### 2018-05-15 WholePlotTrajectories.Rmd
Results & Appendix

To counter the remaining taxonomic uncertainty, plots taxonomic composition and diversity were analysed at the genus level, *i.e.* referring to the genus of observed or trialed botanical names.
Variations in plots taxonomic and functional composition after disturbance was visualized by their trajectories in a two-dimensional ordination space over 30 years.
Two NMDS were conducted, either from taxonomic flora inventories or from plots functional composition based on the 8 leaf, stem and life history traits. 
Plots trajectories along time was reported comparatively to the inventories in 1989, 5 years after disturbance, which corresponded to first inventory with a sufficient degree of uncertainty (<30% of undetermined trees). 
The inventories dissimilarity compared to the reference 1989 inventory was reported using occurrence-based (Jaccard) and abundance-based (Bray-Curtis) similarity measures. 
The trajectory of inventories along time was visualized with the euclidean distance in the two-dimensional ordination space to the 1989 inventory.
The trajectories of the leaf and stem and life traits were also visualized with the community weighted means (CWM), representing the average trait value in a community weighted by relative abundance of the species carrying each value [@Diaz2007;@Garnier2004]. 
To compensate the intrinsinc difference among plots the trajectories corresponded to the differences along time with the reference inventory in 1989.
Species seed mass corresponded to 5 classes of increasing mass, seed mass trajectories were therefore reported as the proportion of each class in the inventories.

@@

## Disturbance impacts on communities composition and average functional traits values

### Composition trajectories

Over time, 828388 trees and 591 botanical species pertaining to 223 genus and 64 botanical families were recorded. 
Taxonomic and functional plots trajectories were examined through the trajecotries in a two dimensional ordination space of successive inventories from 1989 (5 years after disturbance) to 2015 (31 years after disturbance).
Classifications were performed using either abundance-based Bray-Curtis (Figure \@ref(fig:NMDSplans)) or incidence-based Jaccard dissimilarity, this last having given similar results (data not shown).

Both taxonomic and functional composition of disturbed plots followed consistent trajectories over time, translating compositional changes while control plots remained stable.
The addition of functional traits in the two dimensional space revealed a shift of plots composition towards more acquisitive functional strategies (disturbed plots changed from high average WD to high average SLA and chlorophyll content).
The dissimilarity of the successive inventories for disturbed plots compared to their respective reference inventory in 1989 followed unimodal trajectories (AppendixI, Figure A2).
The maximum dissimilarity of plot strajectories was positively correlated to the disturbance intensity for taxonomic ($\rho_{spearman}^{taxonomic}=0.91$) as well as functional composition ($\rho_{spearman}^{taxonomic}=0.96$). 
The time at maximum dissimilarity for taxonomic composition was reached around 26 years after disturbance while it was 22 years for functional composition. 
All trajectories returned towards the initial composition and translated cyclic compositional changes as visualized in figure (Figure \@ref(fig:NMDSplans)).

### Traits community weighted means

Except for leaf cholorphyll content, which continued to increase for some T3 and T2 plots 30 years after disturbance, all traits and seed mass proportions displayed a unimodal trajectories.
The weighted means of communities specific maximum height at adult stage (*Hmax*), leaf toughness (*L_toughness*) and wood specific gravity (*WD*) remained significantly lower than their initial value and than these of the control plots (Figure \@ref(fig:CWM)).
The weighted means of bark thickness (*Bark_thick*) similarly remained substantilly higher than initially for all disturbed plots while the specfic leaf area (*SLA*) had almost recovered its initial value and this of the undisturbed plots at the end of the experiment.
Once again the maximum difference to initial state was correlated to the disturbance intensity ($\rho_{spearman}^{L_{thickness}}=0.67$, $\rho_{spearman}^{L_{chloro}}=0.45$, $\rho_{spearman}^{L_{toughness}}=-0.43$, $\rho_{spearman}^{SLA}=0.93$, $\rho_{spearman}^{WD}=-0.78$, $\rho_{spearman}^{Bark-thickness}=0.88$, $\rho_{spearman}^{Hmax}=-0.48$).

## Trajectories of communities diversity after disturbance

### Communities taxonomic and functional diversity

Trajectories of Richness, Shannon and Simpson taxonomic diversity were examined at genus level in relation to the 1989 inventories (5 years after disturbance) (Figure \@ref(fig:DivTaxo)).

```{r DivTaxo, out.width = '100%', fig.asp = 1.5/3, echo=FALSE, fig.cap="Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities (ref:Bolda) Richness, (ref:Boldb) Shannon and (ref:Boldc) Simpson diversities. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
load("DB/ReplacementTraj_ForGraphs")

TaxoTraj(CompleteTaxo)

For undisturbed plots the Richness, Shannon and Simpson diversities remained stable over the 30 years monitored.
After low disturbance intensity the richness increased, reaching a maximum gain of 14 botanical genera (plot 3 from treatment 2).
After intense disturbance however, the taxonomic richness followed a unimodal decrease with a return to pre-disturbance values.
The taxonomic evenness (Shannon and Simpson diversities) increased after disturbance, whenever its intensity.
The evenness followed a unimodal trajectory with a just beginning return towards initial values and a maximum value, reached around 20 years, positively correlated to the disturbance intensity ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
Only two T3 plots (plots 8 and 12) were still increasing 30 years after disturbance \@ref(fig:DivTaxo), suggesting a similar be much delayed unimodal trajectory.

Communities functional diversity trajectories were examined through the Rao diversity based on the 7 leaf, stem and life history traits (the seed mass was excluded as a qualitative variable).
The plot 7 from treatment 1 that displayed a constantly outlying diversity was removed from the graphical representation for better readability, see figure A3 in appendix for full graph.
The functional diversity of all undisturbed communities remained stable along the 30 years monitored.
For all disturbed communities the functional diversity followed unimodal trajectories with a marked return towards the initial state (5 years after disturbance) with values comparable to those of control plots.

The impact of disturbance was examined specifically at three times after disturbance (10, 20 and 30 years) through the linear correlation between the disturbance intensity (intial %AGB removed) and the value of communities Simpson and Rao diversities (diversities of order 2) \@ref(fig:IDHplot).
The Simpson diversity was only weakly related to the disturbance intensity ($R^2<0.25$) and only from 20 years after disturbance.
The Rao diversity was more strongly correlated to the disturbance intensity ($0.55<R^2<0.75$) for the three times studied and the impact of disturbance was observed 20 years after disturbance.

@@

The functional redundancy was measured as the sum of weighted overlap among species in communities' functional space.
Functional space were defined in the main two-dimensional space of a PCA analysis based on all leaf and stem and maximum height traits (see appendix I for PCA details). 
Plots functional redundancy compared to the initial value after logging so that only communities dynamics are followed apart from the impact of logging itself.
Communities functional redundancy remained stable in control plots but after disturbance the redundancy trajectories were quite variable (See appendix I).
Globally after most intense disturbance (plots T2 and T3) communities redundancy decreased at first place and then increased to edge, recover or even exceed the initial value.
Considering the functional redundancy restricted to the functional space of the initial inventory, though, all disturbed plots followed similar decreasing humped-shaped trajectories (\@ref(fig:RedFun_rest)).
The maximum redundancy loss observed was positively correlated with the disturbance intensity (XX spearman to be measured) and none of the disturbed plots had recovered the functional redundancy within their initial functional space.

>>>>

To counter the remaining taxonomic uncertainty, plots taxonomic composition and diversity were analysed at the genus level, *i.e.* referring to the genus of observed or trialed botanical names.
Variations in plots taxonomic and functional composition after disturbance was visualized by their trajectories over 30 years in a two-dimensional ordination space.
Two NMDS were conducted, either from taxonomic flora inventories or from plots functional composition based on the 7 leaf, stem and life history traits (without seed mass classes).
In both cases the NMDS were performed using occurrence-based (Jaccard) as well as abundance-based (Bray-Curtis) dissimilarity measures. 
Trajectories along time were reported through the euclidean distance of successive inventories to the reference inventories in 1989, 5 years after disturbance, when the uncertainty degree did not exceed 30% of undetermined trees. 
The trajectories of the leaf and stem and life traits were also visualized with the community weighted means (CWM), representing the average trait value in a community weighted by relative abundance of the species carrying each value [@Diaz2007;@Garnier2004]. 
To compensate the intrinsinc difference among plots the trajectories corresponded to the differences along time with the reference inventory in 1989.
Species seed mass corresponded to 5 classes of increasing mass, seed mass trajectories were therefore reported as the proportion of each class in the inventories.

@@

## Communities Diversity

Communities taxonomic diversity trajectories were examined through the Richness, Shannon and Simpson diversities at genus level, in relation to the 1989 inventories (5 years after disturbance) (See annexe I). 
For undisturbed plots the Richness, Shannon and Simpson diversities remained stable over the 30 years monitored.
After low disturbance intensity the richness increased, reaching a maximum gain of 14 botanical genera (plot 3 from treatment 2).
After intense disturbance, plots taxonomic richness followed unimodal trajectories first decrease for around ten years and then increasing to return to pre-disturbance values.
For all disturbed plots the taxonomic evenness (Shannon and Simpson diversities) increased, following unimodal trajectories with a just beginning return towards initial values after 30 years.
The maximum evenness, reached after around 20 years, was positively correlated to the disturbance intensity ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
Only two T3 plots (plots 8 and 12) were still increasing 30 years after disturbance \@ref(fig:IDHplot), suggesting similar but delayed trajectories.

Communities functional diversity trajectories were examined through the Rao diversity based on the 7 leaf, stem and life history traits (the seed mass was excluded as a qualitative variable).
The plot 7 from treatment 1 displayed a constantly outlying diversity and was removed from the graphical representation for better readability (see appendix for full graphs).
The functional diversity of all undisturbed communities remained stable along the 30 years.
For all disturbed communities the functional diversity followed unimodal trajectories with a return towards the initial state, comparable to the control plots.

The impact of disturbance was examined specifically through the linear correlation between the disturbance intensity (intial %AGB removed) and the Simpson and Rao diversities (diversities of order 2) at three focus times after disturbance (10, 20 and 30 years) \@ref(fig:IDHplot).
The Simpson diversity was weakly related to the disturbance intensity ($R^2<0.25$) and only from 20 years after disturbance.
The Rao diversity was more strongly correlated to the disturbance intensity ($0.55<R^2<0.75$) for the three times studied.
The impact of disturbance, translated by the slope of the linear correlation, was the highest 20 years after disturbance.

@@

## Communities composition

### Taxonomic and functional trajectories

In the inventories from 1989 (5 years after disturbance) to 2015 (31 years after disturbance), 828388 trees and 591 botanical species pertaining to 223 genus and 64 botanical families were recorded. 
The trajectories of taxonomic and functional composition were visualized in a two dimensional ordination space mapping the successive inventories according to their flora and the corresponding traits.
Classifications were performed using either abundance-based Bray-Curtis (Figure \@ref(fig:NMDSplans)) or incidence-based Jaccard dissimilarity.
Both metrics gave similar results, so only analysis using Bray-Curtis dissimilarity will be discussed here.

While both taxonomic and functional composition remained stable in undistrubed communities, after disturbed they followed consistent trajectories over time that revealed significant compositional changes.
Considering the mapping of functional traits in the same dimensional space, the post-disturbance trajectories corresponded to shifts towards more acquisitive functional strategies: disturbed communities changed from high average WD to high average SLA and chlorophyll content.
For disturbed communities the distance of successive inventories to the reference inventory in 1989 followed unimodal trajectories (AppendixI, Figure A2).
The maximum of communities dissimilarity to their initial state was positively correlated to the disturbance intensity for both taxonomic and functional composition ($\rho_{spearman}^{taxonomic}=0.91$ and $\rho_{spearman}^{functional}=0.96$ respectively). 
The time at maximum dissimilarity was reached around 26 years after disturbance for taxonomic composition and 22 years for functional composition. 
All trajectories followed cyclic composition changes with a return towards the initial composition (Figure \@ref(fig:NMDSplans)).

### Traits community weighted means (CWM)

Except for leaf chlorophyll content, which continued to increase for some T3 and T2 plots, all traits and seed mass proportions followed a unimodal trajectories and either stabilized or returned towards their initial values 30 years after disturbance.
At that time the weighted means of communities specific maximum height at adult stage (*Hmax*), leaf toughness (*L_toughness*) and wood specific gravity (*WD*) remained significantly lower than their initial value and than these of the control plots (Figure \@ref(fig:CWM)).
The weighted means of bark thickness (*Bark_thick*) similarly remained substantilly higher than initially for all disturbed plots while the specfic leaf area (*SLA*) had almost recovered its initial value, similar as those of undisturbed plots.
For all traits the maximum difference to initial state was correlated to the disturbance intensity ($\rho_{spearman}^{L_{thickness}}=0.67$, $\rho_{spearman}^{L_{chloro}}=0.45$, $\rho_{spearman}^{L_{toughness}}=-0.43$, $\rho_{spearman}^{SLA}=0.93$, $\rho_{spearman}^{WD}=-0.78$, $\rho_{spearman}^{Bark-thickness}=0.88$, $\rho_{spearman}^{Hmax}=-0.48$).

### Functional redundancy

Communities functional redundancy was measured as the sum of weighted overlap among species in communities' functional space.
Functional spaces were defined in the main two-dimensional space of a PCA analysis summarizing the 7 leaf, stem, and maximum height traits (see appendix I for PCA details). 
Communities functional redundancy remained stable in control plots but after disturbance the redundancy trajectories were quite variable (See appendix I).
Globally after most intense disturbance (plots T2 and T3) communities redundancy decreased at first place before increasing to edge, recover or exceed the initial value.

Considering the functional redundancy restricted to the functional space of the initial inventory, all disturbed plots followed similar decreasing humped-shaped trajectories (\@ref(fig:RedFun_rest)).
The maximum redundancy loss observed was positively correlated with the disturbance intensity (XX spearman to be measured) and none of the disturbed plots had recovered the functional redundancy within the initial functional space.
